### 2016-05-20 thesis/thesis.tex
added glossary file

\documentclass[a4paper, 12pt, oneside]{report}

% Preamble:
\usepackage[T1]{fontenc}                     % package for specifying output font encoding
\usepackage[utf8]{inputenc}                  % package for specifying input font encoding
\usepackage{natbib}                          % package for reference management
\usepackage{graphicx}                        % package for graphics handling (i.e. figure images)
\usepackage{mathptmx}                        % package for Times New Roman font
\usepackage{chngcntr}                        % package for making continuous figure numbering, instead of section-based numbering
\usepackage{setspace}                        % package for setting line space (used in title page)
\usepackage[hidelinks, colorlinks]{hyperref} % package for hyperlinking references - hidelinks option removes the ugly boxes around the huperlinks
\usepackage[md]{titlesec}                    % package for tweaking the chapter/section/subsection titles format
\usepackage[acronym, toc]{glossaries}          % package for making glossaries

% Set up the hyperlinks - hange the colour of the links to blue:
\hypersetup{allcolors=blue}

% change figure numbering from section-based to continuous:
\counterwithout{figure}{chapter}
\counterwithout{equation}{chapter}
\counterwithout{table}{chapter}

% If you want to compile specific files, uncomment the line below and add the filenames in the {} :
%\includeonly{}

% Change the title of bibliography to references:
\renewcommand{\bibname}{References}

% To generate glossaries:
\include{./misc/glossary}
\makeglossaries

\begin{document}

\include{./misc/title}

\pagenumbering{roman} % set page numbering to roman numeral

\include{./misc/abstract}

\include{./misc/acknowledgement}

\tableofcontents

\addcontentsline{toc}{section}{\contentsname}

\clearpage

\addcontentsline{toc}{section}{\listfigurename}

\listoffigures

\clearpage

\addcontentsline{toc}{section}{\listtablename}

\listoftables

\clearpage

\printglossaries

\clearpage

% \listoftables

% \clearpage

\pagenumbering{arabic} % revert the page numbering to normal numbering
\glsresetall

\include{./intro/intro}

\include{./methods/methods}

\include{./results/results}

\include{./results/results2}

\include{./discussion/discussion}

% Use the reference style specified in BiocRefStyle.bst:
\bibliographystyle{BiocRefStyle}

% Print out the bibliography:
\bibliography{MSc}

\end{document}