### 2016-06-14 thesis/methods/methods.tmp.tex
re-writing methods section (it was too clunky)

\chapter{Methods (draft)}
\label{ch:methods}

\section{\gls{bmi}}
\label{sec:bmi}

\subsection{\gls{bmi} Calculation}
\label{subsec:bmicalc}

\gls{bmi} was calculated from the height and weight data of the samples using the following equation:

\begin{equation}
	\label{eq:bmicalc}
	\gls{bmi} = \frac{Weight (kg)}{Height^2(m^2)}\\
\end{equation}

\subsection{\gls{bmi} Classification}
\label{subsec:bmiclassification}

Samples were classified based on the \gls{who} definition, as shown in Table~\cref{tab:whobmiclass}.
\begin{table}[hb]
	\caption{\gls{who} defined \gls{bmi} classification}
	\label{tab:whobmiclass}
	\begin{center}
		\begin{tabular}{lc}
			\bfseries {Classification} & \bfseries {\gls{bmi} Value}\\
			\hline
			\rule{0pt}{2.25ex}Underweight & \textless{} 20.0\\
			Normal weight/lean & 20.0$\sim$24.9\\
			Overweight & 25.0$\sim$29.9\\
			Obese & \textgreater{} 30.0\\
		\end{tabular}
	\end{center}
\end{table}

\section{Publicly available cancer data}
\label{sec:data}

Data related to multiple different cancer types were downloaded from free and publicly available sources such as \gls{tcga} and the \gls{icgc}.

\subsection{\gls{tcga}}
\label{subsec:tcga}

The clinical data for all available cancer types (33 types in total) were downloaded from \gls{tcga} database and were checked for the height and weight data for each sample, for each cancer type.
Any cancer type with no weight and/or height data of the samples were excluded from the project, as no \gls{bmi} information can be obtained without these data.

This gave a total of eight cancer types: \gls{blca}, \gls{cesc}, \gls{coad}, \gls{kirp}, \gls{lihc}, \gls{read}, \gls{skcm}, \gls{ucec}.
All eight cancer types were used for further analyses.

\subsection{\gls{icgc}}
\label{subsec:icgc}

\gls{rnaseq} data for the eight cancer types with sample \gls{bmi} information available were downloaded from the \gls{icgc} database.
The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}.
%TODO: add in how I processed the ICGC data format into genes by samples.

%TODO: change the title name?
\subsection{\gls{icgc} sample exclusion criteria}
\label{subsec:sampleexclusion}

Samples from the downloaded data were removed using the following criteria.

\subsubsection{\textbf{Sample BMI availability}}
\label{subsubsec:samplebmiavailability}

Some of the samples did not have either height or weight data, or both.
These samples were removed from the analyses.

\subsubsection{\textbf{Sample ID conversion and matching}}
\label{subsubsec:sampleidconversion}

%TODO: check the actual ID stripping process
%TODO: check the WGCNA function and check if it uses median/mean/whatever for choosing the sample

\gls{icgc} sample ID in the \gls{rnaseq} data were stripped so that the ID from the \gls{rnaseq} data matched with the \gls{tcga} ID in the clinical data.
After stripping the ID, samples were checked for any overlaps.
Where there was an overlap in the sample ID in the \gls{rnaseq} data, a single sample was chosen to represent that particular ID by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.

In each of the eight cancer types, the \gls{tcga} sample ID in the clinical data were cross-checked with the \gls{tcga} sample ID in the \gls{rnaseq} data, and vice versa.
Any sample that did not have either clinical or \gls{rnaseq} data were removed, so both clinical and \gls{rnaseq} data were available for all of the samples that were included in the analyses.
See \cref{tab:samplesize} for a summary of the number of samples included in the analyses for each cancer type.

\begin{table}[h]
	\caption{Summary of the total number of samples included in the analyses for each cancer type}
	\label{tab:samplesize}
	\begin{center}
		\begin{tabular}{cc}
			\textbf{Cancer Type}   & \textbf{Total} \\
			\hline
			\rule{0pt}{2.25ex}BLCA & 261   \\
			CESC                   & 224   \\
			COAD                   & 226   \\
			KIRP                   & 124   \\
			LIHC                   & 264   \\
			READ                   & 73    \\
			SKCM                   & 218   \\
			UCEC                   & 482   \\
		\end{tabular}
	\end{center}
\end{table}

\section{Data processing -- Creighton data}
\label{sec:dataprocessingcr}

\subsection{Raw data from the \citet{Creighton2012} study}
\label{subsec:rawdatacr}

The raw microarray gene expression data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
Clinical data of the samples were obtained from the supplementary table 1 from their paper.
Raw data were imported into \gls{r}/RStudio for subsequent analyses.

%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form

\subsection{Normalisation of raw Creighton data}
\label{subsec:normcrdata}

The raw microarray data for all 103 samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
This ensured that all of the genes were on the same scale as one another, allowing for comparison.

\subsection{Obesity associated gene probes from the \citet{Creighton2012} study}
\label{subsec:crobsgene}

799 obesity associated gene probes identified by \citet{Creighton2012} were obtained from the supplementary data file 1 from their paper.
The list of gene probes were imported into \gls{r}.
The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Any duplicates and any gene symbols that were not present in both the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.

\section{Validation of results -- \citet{Creighton2012}}
\label{sec:valresultscr}

The results presented by \citet{Creighton2012} were validated using the same raw data they have used in their study.
See  \cref{sec:dataprocessingcr} for details on how the raw data was accessed and processed.

\subsection{Creighton metagene generation}
\label{subsec:crmetagen}

The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
All of the *** obesity associated genes from  \cref{subsec:crobsgene} were selected from the data.

\gls{svd} was applied to this data to obtain the decomposed matrices of the data.
%TODO: how do I explain how I got the \gls{metagene} scores?
The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the (orthogonal matrix...?).
The \gls{metagene} scores were ranked based on the number of samples in the data (103 samples in the Creighton data), and the values of the scores were between 0 and 1.

\subsection{Metagene and sample gene expression pattern}
\label{subsec:crmetage}

\Gls{metagene} scores from \cref{subsec:crmetagen} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Metagene and sample \gls{bmi}}
\label{subsec:crmetabmi}

The same \gls{metagene} scores as in the previous section (\cref{subsec:crmetage}) were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.

%TODO: check the formal definition of these test/values
For the boxplot, the p-value was obtained from the ANOVA test.
For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.

\section{Creighton metagene in \gls{icgc} cancer data}
\label{sec:crmetaicgc}

\subsection{\gls{icgc} data processing}
\label{subsec:datprocicgc}

The obesity associated genes from \cref{subsec:crobsgene} were selected in each of the eight cancer types.
All cancer data had 1 added to the count data, so no infinite values were introduced when the data was logged.
All cancer data were logged to the base 10.

All of the data were then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
Any ``NaN'' values in the data introduced from this were assigned a value of 0.

\subsection{Transformation matrix}
\label{subsec:transmatcr}

Transformation matrix was formed from the decomposed matrices that resulted from the application of \gls{svd} to the Creighton data (see \cref{subsec:crmetage}).
%TODO: add the derivation of transformation matrix??

\subsection{Transformed \gls{icgc} data and metagene}
\label{subsec:icgctranscr}

Each of the eight cancer data from \cref{subsec:datprocicgc} were transformed using the transformation matrix from \cref{subsec:transmatcr}.
Metagene scores of the samples in each cancer types were obtained by taking the first eigenvector of the cancer matrix that resulted from the transformation.
The \gls{metagene} scores were ranked based on the number of samples in the data, and the scores ranged from 0 to 1.

\subsection{Creighton metagene and sample gene expression}
\label{subsec:crmetaicgcge}

\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes on a heatmap for each cancer type.
The data from \cref{subsec:datprocicgc} were used for the heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Creighton metagene and sample \gls{bmi}}
\label{subsec:crmetaicgcge}

The same \gls{metagene} scores as in the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values for each cancer type.
Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.

%TODO: check the formal definition of these test/values
For the boxplot, the p-value was obtained from the ANOVA test.
For the scatter plot, the regression line was plotted and the adjusted $R^2$ value was calculated for the regression line.

%TODO: Not sure if I should cut this section to "see previous section for details"
\section{Data processing -- Fuentes-Mattei data}
\label{sec:dataprocessingfm}

\subsection{Raw data from the \citet{Fuentes-Mattei2014} study}
\label{subsec:rawdatafm}

%TODO: add data/sample composition -- ethinicity, grade, etc -- maybe in table form

\subsection{Normalisation of raw Fuentes-Mattei data}
\label{subsec:normfmdata}

The raw microarray data for all *** samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.

\subsection{Obesity associated gene probes from the \citet{Fuentes-Mattei2014} study}
\label{subsec:fmobsgene}

Any gene symbols that were not present in the Fuentes-Mattei data, the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.

\section{Validation of results -- \citet{Fuentes-Mattei2014}}
\label{sec:valresultsfm}

The results presented by \citet{Fuentes-Mattei2014} were validated using the Creighton data, as the clinical data of the samples were not available.
See \cref{subsec:rawdatacr,subsec:normcrdata} for details on how the data was processed.

\subsection{Fuentes-Mattei transformation matrix}
\label{subsec:transmatfm}

\gls{svd} was applied to Fuentes-Mattei data to obtain the decomposed matrices of the data, and this was used to make the transformation matrix.
%TODO: add the derivation of transformation matrix??

\subsection{Fuentes-Mattei metagene generation}
\label{subsec:fmmetagen}

The gene probe IDs in the Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was a duplicated gene, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
All of the *** obesity associated genes from  \cref{subsec:fmobsgene} were selected from the data.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.

%TODO: how do I explain how I got the \gls{metagene} scores?
The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the transformed Creighton data.
The \gls{metagene} scores were ranked based on the number of samples in the data (*** samples in the Creighton data), and the scores ranged from 0 to 1.

\subsection{Fuentes-Mattei metagene and gene expression in the Creighton data}
\label{subsec:fmmetage}

These \gls{metagene} scores were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Fuentes-Mattei metagene and sample \gls{bmi} in the Creighton data}
\label{subsec:fmmetabmi}

The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

\section{Fuentes-Mattei metagene in \gls{icgc} cancer data}
\label{sec:fmmetaicgc}

\subsection{\gls{icgc} data processing}
\label{subsec:datprocicgcfm}

All of the \gls{icgc} data were processed in a similar fashion as in \cref{subsec:datprocicgc}, but the obesity associated genes from \cref{subsec:fmobsgene} were used instead.

\subsection{Transformation matrix}
\label{subsec:icgctransmatfm}

The same transformation matrix from \cref{subsec:transmatfm} was used.

\subsection{Transformed \gls{icgc} data and metagene}
\label{subsec:icgctransfm}

Each of the eight cancer data from \cref{subsec:datprocicgcfm} were transformed using the transformation matrix from \cref{subsec:transmatfm}.
Metagene scores were ranked as in \cref{subsec:icgctranscr}.

\subsection{Fuentes-Mattei metagene and sample gene expression}
\label{subsec:fmmetaicgcge}

\Gls{metagene} scores from \cref{subsec:icgctransfm} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
The data from \cref{subsec:datprocicgcfm} were used for the heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Creighton metagene and sample \gls{bmi}}
\label{subsec:fmmetaicgcge}

The same \gls{metagene} scores as in the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values for each cancer type.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

\section{Differential gene expression analysis}
\label{sec:deg}

\subsection{Data processing and subsetting}
\label{subsec:subdatadeg}

Raw data from \cref{subsec:rawdatacr,subsec:normcrdata} were used in the analysis.
Together with the raw data, few other data sets were created based on the raw data: residual data, Caucasian-only raw data, and Caucasian-only residual data.

Residual data was created from the raw data after it has been controlled for all clinical variables, except those related to \gls{bmi}.
These clinical variables included age, sex, ethinicity, tumour grade, \gls{pr}, \gls{er}, \gls{her2} statuses, and \gls{ln} status.

Caucasian-only data was created by taking only the Caucasian samples from the raw data, which reduced the sample size from 103 samples to 77 samples.
Caucasian-only residual data was created using the same approach as with the residual data: the data that remained after all the non-\gls{bmi} clinical variables were controlled in the Caucasian-only raw data.

\subsection{Group classification}
\label{subsec:crdegclass}

In each of the datasets from \cref{subsec:subdatadeg}, all of the samples were split into two groups: obese and non-obese groups.
These two groups were classified based on the sample \gls{bmi}, where the obese groups were defined as samples with \gls{bmi}  \textgreater{} 30, and non-obese group were defined as samples with \gls{bmi}  \textless{} 30 (see \cref{tab:whobmiclass}).

\subsection{Gene expression analysis}
\label{subsec:crdeg}

The difference in the gene expression between the obese and non-obese sample groups in each of the four data sets described in \cref{subsec:subdatadeg} were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
From the list of significantly expressed (genes produced by the \texttt{topTable} function), all of the gene probes that had a p-value \textless{} 0.01 were selected.

When there was more than 799 gene probes in the list of significant genes, the top 799 gene probes were selected from the list.
This list of 799 gene probes represented the gene probes that were associated with obesity  for that dataset.

\subsection{Common gene probes from the \citet{Creighton2012} study}
\label{subsec:crol}

For each gene probe set, the gene probes that were common with the obesity associated genes identified by \citet{Creighton2012} were selected and used as a separate gene probe set.
There were eight gene probe sets in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from the \citet{Creighton2012} study.

%TODO: add a table summarizing how many genes were in each gene probe set

\subsection{Gene probe set conversion and selection}
\label{subsec:gpconv}

For each of the eight gene probe sets identified in \cref{subsec:crdeg,subsec:crol}, the gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Any duplicates and any gene symbols that were not present in both the Creighton data and the ICGC data were removed from the gene probe set.

\section{BMI metagenes in the Creighton data}
\label{sec:bmidegmetacr}

\subsection{BMI metagene generation}
\label{subsec:bmimetagen}

The gene probe IDs in the Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was a duplicated gene, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.
For each gene probe sets identified in \cref{subsec:crdeg,subsec:crol}, the gene probes were selected from the Creighton data.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.

SVD was applied to the standardised data, and the first eigenvector of the decomposed matrix was used as the sample metagene score for that gene probe set.
The \gls{metagene} scores were ranked based on the number of samples in the data (103 samples in the Creighton data), and the scores ranged from 0 to 1.

\subsection{BMI metagenes and sample gene expression}
\label{subsec:bmimetavscr}

For each \gls{metagene} scores in \cref{subsec:bmimetagen} were plotted together with the gene expression of the associated gene probes for each sample on a heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{BMI metagenes and sample \gls{bmi}}
\label{subsec:bmidegmetacr}

The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

\section{BMI metagenes in the \gls{icgc} data}
\label{sec:bmidegmetaicgc}

\subsection{\gls{icgc} data processing}

All of the \gls{icgc} data were processed in a similar fashion as in \cref{subsec:datprocicgc}, but each of the eight BMI metagene gene probes from \cref{subsec:bmimetagen} were used instead.

\subsection{BMI metagene transformation matrix}
\label{subsec:bmimetagenicgc}

Transformation matrix for each BMI metagene gene probe sets were created using the decomposed matrices of the Creighton data that resulted from the application of SVD in \ref{subsec:bmimetagen}.

\subsection{BMI metagenes and sample gene expression in the \gls{icgc} data}
\label{subsec:gpconv}

For each \gls{metagene} scores in \cref{subsec:bmimetagen} were plotted together with the gene expression of the associated gene probes for each sample on a heatmap for each cancer type.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{BMI metagenes in \gls{icgc} data}
\label{subsec:bmidegmetaicgc}

The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

\newpage

%TODO: need svd derivation and why I'm using it

%TODO: How you could transfer the metagene to other cancer data

%TODO: why transformation matrix is used
%TODO: refer back to eqiations in the  svd section

\section{Gene expression analysis in \gls{icgc} data}
\label{sec:degicgc}

\section{Pathway enrichment analysis in \gls{icgc} data}
\label{sec:pathenrich}