### 2016-11-19 thesis/thesis.tex

% \documentclass[a4paper, 12pt, oneside]{report} % for final draft compilation
\documentclass[a4paper, 12pt, oneside, draft]{report} % for draft compilation

% Preamble:
\usepackage[T1]{fontenc}                            % package for specifying output font encoding
\usepackage[utf8]{inputenc}                         % package for specifying input font encoding
\usepackage{natbib}                                 % package for reference management
\usepackage{graphicx}                               % package for graphics handling (i.e. figure images)
\usepackage{mathptmx}                               % package for Times New Roman font
\usepackage{chngcntr}                               % package for making continuous figure numbering, instead of section-based numbering
\usepackage{setspace}                               % package for setting line space (used in title page)
\usepackage[hidelinks, colorlinks]{hyperref}        % package for hyperlinking references - hidelinks option removes the ugly boxes around the huperlinks
\usepackage[md]{titlesec}                           % package for tweaking the chapter/section/subsection titles format
\usepackage[acronym, toc]{glossaries}               % package for making glossaries
\usepackage{xcolor,colortbl}                        % package for making colours
\usepackage[noabbrev,capitalise]{cleveref}          % package for referencing multiple referencing in a single \ref command
\usepackage[font=footnotesize,format=hang]{caption} % package for formatting figure caption

% Define colour(s) for table rows:
\definecolor{LightCyan}{rgb}{0.88,1,1}

% Set up the hyperlinks - change the colour of the links to blue:
% \hypersetup{allcolors=blue}
\hypersetup{allcolors=black} % uncomment this line to make the hyperlinks black

% Change the default cref name for equations:
\creflabelformat{equation}{#2\textup{#1}#3}

% Change the default url font:
\urlstyle{rm}

% change figure numbering from section-based to continuous:
\counterwithout{figure}{chapter}
\counterwithout{equation}{chapter}
\counterwithout{table}{chapter}

% Change the title of bibliography to references:
\renewcommand{\bibname}{References}

% Italicise the subsubsection title:
\titleformat{\subsubsection}{\normalfont\itshape\bfseries}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}{\normalfont\itshape}{\theparagraph}{\noindent}{}

% For generating glossaries:
\include{./misc/glossary}
\makeglossaries

% Specify graphics path:
\graphicspath{ {./images/}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

% If you want to compile specific files, uncomment the line below and add the filenames in the {} :
%\includeonly{}

\begin{document}

\include{./misc/title}

\pagenumbering{roman} % set page numbering to roman numeral

% Set line spacing to double space:
\onehalfspacing

\include{./misc/abstract}

\include{./misc/acknowledgement}

\tableofcontents

\addcontentsline{toc}{section}{\contentsname}

\clearpage

\addcontentsline{toc}{section}{\listfigurename}

\listoffigures

\clearpage

\addcontentsline{toc}{section}{\listtablename}

\listoftables

\clearpage

\printglossaries

\clearpage

% \listoftables

% \clearpage

\pagenumbering{arabic} % revert the page numbering to normal numbering
\glsresetall

\include{./intro/intro}

\include{./methods/methods}

\include{./results/results}

\include{./results/results2}

\include{./discussion/discussion}

% Use the reference style specified in BiocRefStyle.bst:
\bibliographystyle{BiocRefStyle}

% Print out the bibliography:
\bibliography{MSc}

\end{document}