### 2017-01-06 thesis/misc/glossary.tex
Finished draft copy of introduction -- need to cut down on the contents and link everything together

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Glossary terms
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% TODO: make sure you get a proper definitions for all of these glossary terms

\newglossaryentry{rnaseq}{
	name=RNA-seq,
	description={sequence data using transcriptome}
}

\newglossaryentry{r}{
	name=R,
	description={Statistical programming language}
}

\newglossaryentry{metagene}{
	name=Metagene,
	text=metagene,
	description={Summary score of a sample that represents the overall gene expression of the genes that was used to obtain the scores}
}

\newglossaryentry{ldp}{
	name=Limited duration prevalence,
	text=limited duration prevalence,
	description={Number of patients diagnosed with a disease within a fixed period of time in the past \citep{Bray2013}}
}

\newglossaryentry{apoptosis}{
	name=Apoptosis,
	text=apoptosis,
	description={The death of cells which occurs as a normal and controlled part of an organism's growth or development}
}

\newglossaryentry{chromothripsis}{
	name=Chromothripsis,
	text=chromothripsis,
	description={An all-at-once mutational pattern in which a large number of rearrangements are confined to local regions of one or more chromosomes \citep{Stephens2011}}
}

\newglossaryentry{kataegis}{
	name=Kataegis,
	text=kataegis,
	description={Showers of single nucleotide changes confined to local regions, often near sites of chromosome rearrangements \citep{Stephens2011}}
}

\newglossaryentry{adipocyte}{
	name=Adipocyte,
	text=adipo\-cyte,
	plural=adipo\-cytes,
	description={Adipose tissue cell}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Acronym terms
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newacronym{blca}{BLCA}{bladder urothelial carcinoma}
\newacronym{cesc}{CESC}{cervical squamous cell carcinoma and endocervical adenocarcinoma}
\newacronym{coad}{COAD}{colon adenocarcinoma}
\newacronym{kirp}{KIRP}{kidney renal clear cell carcinoma}
\newacronym{lihc}{LIHC}{liver hepatocellular carcinoma}
\newacronym{read}{READ}{rectum adenocarcinoma}
\newacronym{skcm}{SKCM}{skin cutaneous melanoma}
\newacronym{ucec}{UCEC}{uterine corpus endometrial carcinoma}
\newacronym{geo}{GEO}{Gene Expression Omnibus}
\newacronym{tcga}{TCGA}{The Cancer Genome Atlas}
\newacronym{icgc}{ICGC}{International Cancer Genome Consortium}
\newacronym{bmi}{BMI}{Body Mass Index}
\newacronym{who}{WHO}{World Health Organization}
\newacronym{ln}{LN}{lymph node}
\newacronym{pr}{PR}{progesterone receptor}
\newacronym{er}{ER}{estrogen receptor}
\newacronym{her2}{HER2}{human epidermal growth factor receptor 2}
\newacronym{svd}{SVD}{singular value decomposition}
\newacronym{rma}{RMA}{Robust Multi Array}
\newacronym{mas}{MAS5}{Microarray Analysis Suite 5.0}
\newacronym{fdr}{FDR}{False Discovery Rate}
\newacronym{go}{GO}{Gene Ontology}
\newacronym{usa}{USA}{United States of America}
\newacronym{npy}{NPY}{neuropeptide Y}
\newacronym{agrp}{AGRP}{agouti-related protein}
\newacronym{pomc}{POMC}{pro-opiomelanocortin}
\newacronym{cart}{CART}{cocaine and amphetamine related transcript}
\newacronym{mc1r}{MC1R}{melanocortin 1 receptor}
\newacronym{mc4r}{MC4R}{melanocortin 4 receptor}
\newacronym{msh}{MSH}{melanocyte-stimulating hormone}
\newacronym{nyr1}{Y1R}{neuropeptide Y1 receptor}
\newacronym{acth}{ACTH}{adrenocorticotropic hormone}
\newacronym{ghsr}{GHSR}{growth hormone secretagogue receptor}
\newacronym{gaba}{GABA}{$\gamma$-aminobutyric acid}
\newacronym{irs}{IRS}{insulin receptor seubstrate}
\newacronym{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
\newacronym{jak}{JAK}{janus kinase}
\newacronym{pc1}{PC1}{prohormone convertase 1}
\newacronym{fto}{FTO}{fat mass and obesity associated}
\newacronym{fto}{FTO}{fat mass and obesity associated}
\newacronym{snp}{SNP}{single nucleotide polymorphism}
\newacronym{rb}{RB}{retinoblastoma-associated}
\newacronym{opa}{OPA}{obesity policy action}
\newacronym{g1}{G1}{Gap 1}
\newacronym{bcl2}{Bcl-2}{B-cell lymphoma 2}
\newacronym{bh3}{BH3}{Bcl-2 homology 3}
\newacronym{dna}{DNA}{deoxyribonucleic acid}
\newacronym{vegf}{VEGF}{vascular endothelial growth factor}
\newacronym{vegfa}{VEGF-A}{vascular endothelial growth factor-A}
\newacronym{tsp1}{TSP-1}{thrombospondin-1}
\newacronym{emt}{EMT}{endothelial-mesenchymal transition}
\newacronym{met}{MET}{mesenchymal-endothelial transition}
\newacronym{aid}{AID}{activation-induced cytidine deaminase}
\newacronym{apobec}{APOBEC}{apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}
\newacronym{nfkb}{NF-$\kappa$B}{nuclear factor $\kappa$B}
\newacronym{stat3}{STAT3}{signal transducer and activator of transcription 3}
\newacronym{il6}{IL-6}{interleukin-6}
\newacronym{il10}{IL-10}{interleukin-10}
\newacronym{atp}{ATP}{adenosine triphosphate}
\newacronym{pep}{PEP}{phosphoenol pyruvate}
\newacronym{mmr}{MMR}{mismatch repair}
\newacronym{mtor}{mTOR}{mammalian target of rapamycin}
\newacronym{ner}{NER}{nucleotide-excision repair}
\newacronym{ber}{BER}{base-excision repair}
\newacronym{uv}{UV}{ultraviolet}
\newacronym{pah}{PAH}{polycyclic aromatic hydrocarbon}
\newacronym{nnk}{NNK}{4-(methylnitrosamino)-1-(3-pyridyl)-1-butanone}
\newacronym{ros}{ROS}{reactive oxygen species}
\newacronym{cyp}{CYP}{cytochrome P450}
\newacronym{igf}{IGF}{insulin-like growth factor}
\newacronym{igfbp}{IGFBP}{insulin-like growth factor binding protein}
\newacronym{gpcr}{GPCR}{G-protein coupled receptor}
\newacronym{tlr}{TLR}{Toll-like receptor}
\newacronym{prr}{PRR}{pattern recognition receptor}
\newacronym{ssdna}{ssDNA}{single-stranded DNA}
\newacronym{erk}{ERK}{extracellular signal regulated kinase}
\newacronym{cdna}{cDNA}{complimentary DNA}
\newacronym{mrna}{mRNA}{messenger RNA}
\newacronym{ngs}{NGS}{next generation sequencing}
\newacronym{fc}{FC}{fold change}
\newacronym{ifny}{IFN-$\gamma$}{interferon-$\gamma$}
\newacronym{tgfb}{TGF-$\beta$}{transforming growth factor-$\beta$}
\newacronym[longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
\newacronym{stat}{STAT}{signal transducer and activator of transcription}
\newacronym{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
\newacronym[longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}