### 2017-03-13 thesis/misc/appendix.tex
Need to add linear model prediction scatter plots - tmp commit

\begin{appendices}
	\crefalias{chapter}{appcha}
	\crefalias{section}{appsec}
	\renewcommand{\thesection}{\Alph{chapter}\arabic{section}}

	\chapter{Additional results for \cref{cha:obesity_genetic_signatures_and_cancer}}
	\label{app:a}

	\section{Comparison of the Creighton \textit{et al.} obesity metagene in standardised or non-standardised CR data}
	\label{sec:metagenes_created_from_raw_data_vs_standardised_data}

	\begin{figure}[htpb]
		\centering
		\includegraphics[width=0.45\linewidth,page=1]{appendix/check_raw_vs_std_mg}
		\hfill
		\includegraphics[width=0.45\linewidth,page=2]{appendix/check_raw_vs_std_mg}
		\caption{Scatter plots showing the raw and ranked Creighton \textit{et al.} obesity metagene values generated from standardised or non-standardised CR data set. }
		\label{fig:appendix/check_raw_vs_std}
	\end{figure}

	\section{Comparison of the Creighton \textit{et al.} obesity metagene in standardised or non-standardised \gls{icgc} data}
	\label{sec:comp_cr_raw_std_icgc}

	As shown in \cref{fig:appendix/check_raw_vs_std}, the Creighton \textit{et al.} obesity metagene values generated from the transformation matrices (from standardised or non-standardised CR data) were more consistent (or less variable) in the standardised \gls{icgc} \gls{blca} data, compared with the non-standardised data.
	Results from \cref{fig:appendix/check_raw_vs_std_tm} suggested that there was no obvious difference in the quality of the metagenes generated from standardised or non-standardised TM.
	Only the results from \gls{blca} data set are shown, as the results from the other \gls{icgc} cancer types showed very similar results.

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.45\linewidth,page=1]{appendix/crtcga_raw_vs_std}
		\hfill
		\includegraphics[width=0.45\linewidth,page=2]{appendix/crtcga_raw_vs_std}\\
		\caption{Scatter plots comparing the Creighton \textit{et al.} obesity metagene generated from transformation matrix (from standardised or non-standardised CR data) in non-standardised (left) and standardised (right) \gls{icgc} \gls{blca} data.}
		\label{fig:appendix/check_raw_vs_std}
	\end{figure}

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.45\linewidth,page=3]{appendix/crtcga_raw_vs_std}
		\hfill
		\includegraphics[width=0.45\linewidth,page=4]{appendix/crtcga_raw_vs_std}\\
		\caption{Scatter plots comparing the Creighton \textit{et al.} obesity metagene generated from transformation matrix from non-standardisedright (left) or standardised (right) CR data in non-standardised  and standardised  \gls{icgc} \gls{blca} data.}
		\label{fig:appendix/check_raw_vs_std_tm}
	\end{figure}

	\section{Remainder of the results of Creighton \textit{et al.} obesity metagenes in the \gls{icgc} cancer data sets}
	\label{sec:rest_of_the_cr_icgc_cancer_heatmap_results}

	As shown in \cref{fig:appendix/cr_icgc_heatmap}, the original obesity metagene from the Creighton \textit{et al.} study was reflective of the sample gene expression levels of the obesity associated genetic signature in all of the other \gls{icgc} cancer data sets.
	However, none of the \gls{icgc} cancer data sets (other than the \gls{blca} data set; \cref{fig:crmetaicgc} in \cref{sec:creighton_obesity_metagene}) showed any significant association of the sample \gls{bmi} or \gls{bmi} status with the Or obesity metagene (\cref{fig:appendix/cr_icgc_box_scatter}).
	The heatmaps, box plots and scatter plots for the other CR obesity metagenes in other \gls{icgc} cancer types are not shown, as many of the plots were very similar to the plots from the \gls{blca} data set.
	The statistics from the results of the box and scatter plots in the other \gls{icgc} cancer types were summarised in \cref{tab:degmetacesc,tab:degmetacoad,tab:degmetakirp,tab:degmetalihc,tab:degmetaread,tab:degmetaskcm,tab:degmetaucec}.

	\begin{figure}[htp!]
		\centering
		\includegraphics[width=0.32\linewidth,page=8]{appendix/crtcga_std}\\
		\vspace{1em}
		\includegraphics[width=0.32\linewidth,page=13]{appendix/crtcga_std}
		\includegraphics[width=0.32\linewidth,page=18]{appendix/crtcga_std}
		\includegraphics[width=0.32\linewidth,page=23]{appendix/crtcga_std}\\
		\vspace{1em}
		\includegraphics[width=0.32\linewidth,page=28]{appendix/crtcga_std}
		\includegraphics[width=0.32\linewidth,page=33]{appendix/crtcga_std}
		\includegraphics[width=0.32\linewidth,page=38]{appendix/crtcga_std}\\
		\caption{Heatmaps showing the association of obesity metagene from the \citet{Creighton2012} study with sample gene expression from the other \gls{icgc} data sets.
	Level of expression is represented in the top right histogram, where low and high gene expression were colour-coded with blue and red, respectively.
	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the \gls{icgc} data.
	The obesity associated metagene scores of the samples are shown in a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown to the left of the heatmap. }
		\label{fig:appendix/cr_icgc_heatmap}
	\end{figure}

	\begin{figure}[htpb]
		\centering
		\includegraphics[page=9,width=0.45\linewidth]{appendix/crtcga_std}
		\hfill
		\includegraphics[page=10,width=0.45\linewidth]{appendix/crtcga_std}\\
		\includegraphics[page=14,width=0.45\linewidth]{appendix/crtcga_std}
		\hfill
		\includegraphics[page=15,width=0.45\linewidth]{appendix/crtcga_std}\\
		\caption{Box plots and scatter plots showing the association of obesity metagene from the \citet{Creighton2012} study with the sample \gls{bmi} status  and \gls{bmi} from the other \gls{icgc} data sets, respectively.
	In the box plot, the p-values above the groups represent the statistical significance of the association of the metagene with the overweight or obese group compared with the normal weight group.
	The \gls{anova} p-value shows the statistical significance of the association of the metagene with the sample \gls{bmi} groups.
	In the scatter plot, $R^2$- and p-values describe the adjusted coefficient of determination of the regression line and the statistical significance of the linear model used to draw the regression line, respectively.}
		\label{fig:appendix/cr_icgc_box_scatter}
	\end{figure}

	\begin{figure}[htpb]
		\ContinuedFloat
		\captionsetup{list=off,format=cont}
		\centering
		\includegraphics[page=19,width=0.45\linewidth]{appendix/crtcga_std}
		\hfill
		\includegraphics[page=20,width=0.45\linewidth]{appendix/crtcga_std}\\
		\includegraphics[page=24,width=0.45\linewidth]{appendix/crtcga_std}
		\hfill
		\includegraphics[page=25,width=0.45\linewidth]{appendix/crtcga_std}\\
		\includegraphics[page=29,width=0.45\linewidth]{appendix/crtcga_std}
		\hfill
		\includegraphics[page=30,width=0.45\linewidth]{appendix/crtcga_std}\\
		\caption{Name}
	\end{figure}

	\begin{figure}[htpb]
		\ContinuedFloat
		\captionsetup{list=off,format=cont}
		\centering
		\includegraphics[page=34,width=0.45\linewidth]{appendix/crtcga_std}
		\hfill
		\includegraphics[page=35,width=0.45\linewidth]{appendix/crtcga_std}\\
		\includegraphics[page=39,width=0.45\linewidth]{appendix/crtcga_std}
		\hfill
		\includegraphics[page=40,width=0.45\linewidth]{appendix/crtcga_std}\\
		\caption{Name}
	\end{figure}

\begin{table}[htpb]
	\centering
	\caption{Statistics of all the obesity metagenes in \gls{cesc} cancer data}
	\label{tab:degmetacesc}
	\begin{threeparttable}
		\begin{tabular}{lccccc}
			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
			\cmidrule(r){2-4} \cmidrule(r){5-6}
			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
			\hline
			\hline
			\rule{0pt}{2.25ex}Cr      & 0.9332                      & 0.5302  & 0.7622             & -0.003950  & 0.7265              \\
            Res     & 0.9205                      & 0.3369  & 0.6043             & 0.004249   & 0.8122              \\
            CrOl    & 0.8940                      & 0.1430  & 0.2959             & -0.0004927 & 0.3465              \\
            ResOl   & 0.9930                      & 0.1705  & 0.3078             & 0.003263   & 0.1898              \\
            Ca      & 0.8861                      & 0.3616  & 0.6426             & -0.003307  & 0.6073              \\
            CaRes   & 0.8323                      & 0.3330  & 0.6167             & -0.003832  & 0.7001              \\
            CaOl    & 0.9124                      & 0.1602  & 0.2600             & -0.001612  & 0.4242              \\
            CaResOl & 0.9646                      & 0.1663  & 0.2863             & 0.003097   & 0.1946              \\
			\hline
			\hline
		\end{tabular}
	\end{threeparttable}
\end{table}

\begin{table}[htpb]
	\centering
	\caption{Statistics of all the obesity metagenes in \gls{coad} cancer data}
	\label{tab:degmetacoad}
	\begin{threeparttable}
		\begin{tabular}{lccccc}
			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
			\cmidrule(r){2-4} \cmidrule(r){5-6}
			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
			\hline
			\hline
			\rule{0pt}{2.25ex}Cr      & 0.2843                      & 0.8176  & 0.5460             & -0.004373  & 0.8865              \\
            Res     & 0.4091                      & 0.8473  & 0.7064             & -0.004243  & 0.8245              \\
            CrOl    & 0.4063                      & 0.9295  & 0.6150             & -0.00397   & 0.7400              \\
            ResOl   & 0.3748                      & 0.8484  & 0.5295             & -0.004020  & 0.7532              \\
            Ca      & 0.6507                      & 0.7692  & 0.7579             & -0.002493  & 0.5075              \\
            CaRes   & 0.4565                      & 0.9808  & 0.7161             & -0.004278  & 0.8385              \\
            CaOl    & 0.2931                      & 0.9217  & 0.4621             & -0.00425   & 0.8273              \\
            CaResOl & 0.4203                      & 0.9087  & 0.7017             & -0.004461  & 0.9800              \\
			\hline
			\hline
		\end{tabular}
	\end{threeparttable}
\end{table}

\begin{table}[htpb]
	\centering
	\caption{Statistics of all the obesity metagenes in \gls{kirp} cancer data}
	\label{tab:degmetakirp}
	\begin{threeparttable}
		\begin{tabular}{lccccc}
			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
			\cmidrule(r){2-4} \cmidrule(r){5-6}
			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
			\hline
			\hline
			\rule{0pt}{2.25ex}Cr      & 0.1853                      & 0.2590  & 0.3387             & -0.005657  & 0.5798              \\
            Res     & 0.1063                      & 0.1825  & 0.2128             & -0.005037  & 0.5368              \\
            CrOl    & 0.2576                      & 0.1757  & 0.3162             & -0.005763  & 0.5879              \\
            ResOl   & 0.3120                      & 0.1738  & 0.3198             & -0.004513  & 0.5048              \\
            Ca      & 0.2431                      & 0.2142  & 0.3591             & -0.005582  & 0.5743              \\
            CaRes   & 0.2327                      & 0.2344  & 0.3738             & -0.006250  & 0.6279              \\
            CaOl    & 0.3426                      & 0.2775  & 0.4693             & -0.007333  & 0.77469             \\
            CaResOl & 0.2643                      & 0.1315  & 0.2410             & -0.001408  & 0.3649              \\
			\hline
			\hline
		\end{tabular}
	\end{threeparttable}
\end{table}

\begin{table}[htpb]
	\centering
	\caption{Statistics of all the obesity metagenes in \gls{lihc} cancer data}
	\label{tab:degmetalihc}
	\begin{threeparttable}
		\begin{tabular}{lccccc}
			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
			\cmidrule(r){2-4} \cmidrule(r){5-6}
			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
			\hline
			\hline
			\rule{0pt}{2.25ex}Cr      & 0.4033                      & 0.2900  & 0.5221             & 0.006348   & 0.1028              \\
            Res     & 0.5008                      & 0.8743  & 0.8004             & -0.0003709 & 0.3430              \\
            CrOl    & 0.7599                      & 0.8562  & 0.9189             & -0.001554  & 0.4424              \\
            ResOl   & 0.9238                      & 0.9773  & 0.9938             & -0.001283  & 0.4162              \\
            Ca      & 0.4898                      & 0.8221  & 0.7917             & 0.0005899  & 0.2834              \\
            CaRes   & 0.5976                      & 0.9929  & 0.8638             & -0.001090  & 0.3990              \\
            CaOl    & 0.8717                      & 0.8834  & 0.9679             & -0.002010  & 0.4925              \\
            CaResOl & 0.5369                      & 0.7185  & 0.7073             & -0.0004212 & 0.3465              \\
			\hline
			\hline
		\end{tabular}
	\end{threeparttable}
\end{table}

\begin{table}[htpb]
	\centering
	\caption{Statistics of all the obesity metagenes in \gls{read} cancer data}
	\label{tab:degmetaread}
	\begin{threeparttable}
		\begin{tabular}{lccccc}
			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
			\cmidrule(r){2-4} \cmidrule(r){5-6}
			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
			\hline
			\hline
			\rule{0pt}{2.25ex}Cr      & 0.9129                      & 0.4008  & 0.5942             & -0.0132    & 0.8038              \\
            Res     & 0.7737                      & 0.2244  & 0.4308             & -0.008917  & 0.5484              \\
            CrOl    & 0.8439                      & 0.1283  & 0.2643             & 0.001024   & 0.3036              \\
            ResOl   & 0.6953                      & 0.06841 & 0.1766             & 0.01099    & 0.1840              \\
            Ca      & 0.6298                      & 0.1326  & 0.2960             & -0.001242  & 0.3432              \\
            CaRes   & 0.6302                      & 0.1053  & 0.2453             & 0.001989   & 0.2885              \\
            CaOl    & 0.6871                      & 0.1411  & 0.3358             & -0.004847  & 0.4219              \\
            CaResOl & 0.8821                      & 0.08474 & 0.1817             & 0.001095   & 0.3025              \\
			\hline
			\hline
		\end{tabular}
	\end{threeparttable}
\end{table}

\begin{table}[htpb]
	\centering
	\caption{Statistics of all the obesity metagenes in \gls{skcm} cancer data}
	\label{tab:degmetaskcm}
	\begin{threeparttable}
		\begin{tabular}{lccccc}
			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
			\cmidrule(r){2-4} \cmidrule(r){5-6}
			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
			\hline
			\hline
			\rule{0pt}{2.25ex}Cr      & 0.7115                      & 0.3329  & 0.6264             & 0.002887   & 0.2033              \\
            Res     & 0.7604                      & 0.5408  & 0.6396             & -0.001725  & 0.4296              \\
            CrOl    & 0.4719                      & 0.1224  & 0.2913             & 0.01261    & 0.05347             \\
            ResOl   & 0.8640                      & 0.1668  & 0.3051             & 0.01541    & {\bfseries 0.0372}\tnote{1}  \\
            Ca      & 0.8792                      & 0.2084  & 0.3739             & 0.006674   & 0.1184              \\
            CaRes   & 0.8098                      & 0.1950  & 0.3677             & 0.005746   & 0.1347              \\
            CaOl    & 0.9478                      & 0.2726  & 0.4471             & 0.01009    & 0.07449             \\
            CaResOl & 0.7309                      & 0.2561  & 0.2830             & 0.01606    & {\bfseries 0.03421} \\
			\hline
			\hline
		\end{tabular}
		\begin{tablenotes}
			\item [1] All values in bold are statistically significant (p \textless{} 0.05).
		\end{tablenotes}
	\end{threeparttable}
\end{table}

\begin{table}[htpb]
	\centering
	\caption{Statistics of all the obesity metagenes in \gls{ucec} cancer data}
	\label{tab:degmetaucec}
	\begin{threeparttable}
		\begin{tabular}{lccccc}
			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
			\cmidrule(r){2-4} \cmidrule(r){5-6}
			Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
			\hline
			\hline
			\rule{0pt}{2.25ex}Cr      & 0.8357                      & 0.08117 & 0.1039             & 0.006952   & {\bfseries 0.03716}\tnote{1}\\
            Res     & 0.8664                      & 0.2023  & 0.3131             & 0.001998   & 0.1619              \\
            CrOl    & 0.9251                      & 0.3800  & 0.5800             & -0.000311  & 0.3569              \\
            ResOl   & 0.5826                      & 0.09608 & 0.2156             & 0.003487   & 0.1021              \\
            Ca      & 0.7491                      & 0.1103  & 0.1905             & 0.003811   & 0.09260             \\
            CaRes   & 0.8086                      & 0.2002  & 0.3419             & 0.0005824  & 0.2584              \\
            CaOl    & 0.9068                      & 0.2672  & 0.4153             & 0.00305    & 0.1166              \\
            CaResOl & 0.8360                      & 0.2440  & 0.4034             & 0.00022898 & 0.2863              \\
			\hline
			\hline
		\end{tabular}
		\begin{tablenotes}
			\item [1] All values in bold are statistically significant (p \textless{} 0.05).
		\end{tablenotes}
	\end{threeparttable}
\end{table}

	\section{Remainder of the results of the other CR obesity metagenes in \gls{nzbc} data set}
	\label{sec:rest_of_the_cr_ob_meta_heatmap_results_cris}

	All of the other obesity metagenes from the CR data set were able to capture the overall gene expressions of the samples in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_heatmap_cris}).
	However, none of the CR obesity metagenes were able to significantly associate with the sample \gls{bmi}/\gls{bmi} status in the \gls{nzbc} data set (\cref{fig:appendix/cr_ob_meta_box_scatter_cris}).

	\begin{figure}[htp!]
		\centering
		\includegraphics[width=0.32\linewidth,page=8]{appendix/cris_crdeg_trans_meta}\\
		\includegraphics[width=0.32\linewidth,page=13]{appendix/cris_crdeg_trans_meta}
		\includegraphics[width=0.32\linewidth,page=18]{appendix/cris_crdeg_trans_meta}
		\includegraphics[width=0.32\linewidth,page=23]{appendix/cris_crdeg_trans_meta}\\
		\includegraphics[width=0.32\linewidth,page=28]{appendix/cris_crdeg_trans_meta}
		\includegraphics[width=0.32\linewidth,page=33]{appendix/cris_crdeg_trans_meta}
		\includegraphics[width=0.32\linewidth,page=38]{appendix/cris_crdeg_trans_meta}\\
		\caption{Heatmaps showing the association of the other CR obesity metagenes with the sample gene expression from the \gls{nzbc} data set.
	Scales are as described in previous figures.}
		\label{fig:appendix/cr_ob_meta_heatmap_cris}
	\end{figure}

	\begin{figure}[htpb]
		\centering
		\includegraphics[page=9,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}
		\hfill
		\includegraphics[page=10,width=0.42\linewidth]{appendix/cris_crdeg_trans_meta}\\
		\caption{Box plots and scatter plots showing the association of the other CR obesity metagenes with the sample \gls{bmi} status  and \gls{bmi} from the \gls{nzbc} data set, respectively.
	P-values and $R^2$-value are as described in previous figures.}
		\label{fig:appendix/cr_ob_meta_box_scatter_cris}
	\end{figure}

	\begin{figure}[htpb]
		\ContinuedFloat
		\captionsetup{list=off,format=cont}
		\centering
		\includegraphics[page=14,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
		\hfill
		\includegraphics[page=15,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
		\includegraphics[page=19,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
		\hfill
		\includegraphics[page=20,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
		\includegraphics[page=24,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
		\hfill
		\includegraphics[page=25,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
		\caption{Name}
	\end{figure}

	\begin{figure}[htpb]
		\ContinuedFloat
		\captionsetup{list=off,format=cont}
		\centering
		\includegraphics[page=29,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
		\hfill
		\includegraphics[page=30,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
		\includegraphics[page=34,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
		\hfill
		\includegraphics[page=35,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
		\includegraphics[page=39,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}
		\hfill
		\includegraphics[page=40,width=0.45\linewidth]{appendix/cris_crdeg_trans_meta}\\
		\caption{Name}
	\end{figure}