### 2017-03-29 thesis/misc/acknowledgement.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Acknowledgement for my thesis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\centering Acknowledgement}
\addcontentsline{toc}{section}{Acknowledgement}

First and foremost, I would like to thank my supervisor Mik, for giving me this project and an
opportunity to learn and develop the basic bioinformatics skills required to complete it.
You have been a great supervisor and I have learnt a lot from you over the past two years, and I have enjoyed being a part of the Black lab.
\\

\noindent
I would like to thank my committee members, Dr.\ Anita Dunbier, Dr.\ Miriam Sharpe, and Dr.\ Peter Mace (temporary member), for being patient during my presentations that were full of statistics and heatmaps.
I am grateful for all the suggestions and inputs you have given me to make this a better project.
\\

\noindent
I would also like to thank both past and present members of the Black lab, for useful tips and tricks to get me through my Masters.
Especially James Boocock, for introducing me to a variety of useful tool sets (by the way, everyone should start using vim); Murray Cadzow, for organising the fortnightly ``Sh*t You Should Know About'' sessions to learn more about R and other useful tools; and Tom Kelly, for being my companion in the Fish-bowl and helping me with some of the codes in my project.
\\

\noindent
I would like to thank all of my friends and family for all their support and fun memories throughout the years that I have been at the University.
Lastly, a special thanks to Gezza and Tessa who have generously agreed to proof-read my introduction.
\\