### 2017-03-30 thesis/misc/glossary.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Glossary terms
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newglossaryentry{rnaseq}{
	name=RNA-seq,
	description={Sequence data using the transcriptome}
}

\newglossaryentry{r}{
	name=R,
	description={Statistical programming language}
}

\newglossaryentry{immunoediting}{
	name=Immunoediting,
	text=immunoediting,
	description={Process where the immune system successfully eliminates the highly immunogenic cancer cells during surveillance, but fails to eliminate the weakly immunogenic cells}
}

\newglossaryentry{metagene}{
	name=Metagene,
	text=metagene,
	description={Summary score of a sample that represents the overall gene expression of the genes that was used to obtain the scores}
}

\newglossaryentry{ldp}{
	name=Limited duration prevalence,
	text=limited duration prevalence,
	description={Number of patients diagnosed with a disease within a fixed period of time in the past \citep{Bray2013}}
}

\newglossaryentry{apoptosis}{
	name=Apoptosis,
	text=apoptosis,
	description={The death of cells which occurs as a normal and controlled part of an organism's growth or development}
}

\newglossaryentry{chromothripsis}{
	name=Chromothripsis,
	text=chromothripsis,
	description={An all-at-once mutational pattern in which a large number of rearrangements are confined to local regions of one or more chromosomes \citep{Stephens2011}}
}

\newglossaryentry{kataegis}{
	name=Kataegis,
	text=kataegis,
	description={Showers of single nucleotide changes confined to local regions, often near sites of chromosome rearrangements \citep{Stephens2011}}
}

\newglossaryentry{adipocyte}{
	name=Adipocyte,
	text=adipo\-cyte,
	plural=adipo\-cytes,
	description={Adipose tissue cell}
}

\newglossaryentry{type1}{
	name=Type I error,
	text=Type I error,
	plural=Type I errors,
	description={Incorrect rejection of a true null hypothesis; false positive}
}

\newglossaryentry{type2}{
	name=Type II error,
	text=Type II error,
	plural=Type II errors,
	description={Incorrect retaining of a false null hypothesis; false negative}
}

\newglossaryentry{power}{
	name=Power,
	text=power,
	description={Probability that it will reject a false null hypothesis; it is inversely related to the probability of making a Type II error}
}

\newglossaryentry{ontology}{
	name=Ontology,
	text=ontology,
	description={A set of concepts and categories in a subject area or domain that shows their properties and the relations between them}
}

\newglossaryentry{carcinogen}{
	name=Carcinogen,
	text=carcinogen,
	plural=carcinogens,
	description={A substance capable of causing cancer}
}

\newglossaryentry{passenger}{
	name=Passanger mutation,
	text=passanger,
	description={Mutations that have neither a selective advantage nor contribute to tumour progression}
}

\newglossaryentry{driver}{
	name=Driver mutation,
	text=driver,
	description={Mutations that helps tumorigenesis and its subsequent progression}
}

\newglossaryentry{somatic}{
	name=Somatic mutation,
	text=somatic mutation,
	plural=somatic mutations,
	description={Mutations that are acquried by the cell over the course of the lifetime of an individual}
}

\newglossaryentry{germline}{
	name=Germline mutation,
	text=germline mutation,
	plural=germline mutations,
	description={Mutations that are passed on from the parents}
}

\newglossaryentry{stability}{
	name=Stability gene,
	text=stability gene,
	plural=stability genes,
	description={Genes that are involved in \acrshort{dna} repair mechanisms}
}

\newglossaryentry{allele}{
	name=Allele,
	text=allele,
	plural=alleles,
	description={Variant form of a given gene}
}

\newglossaryentry{suppressor}{
	name=Tumour suppressor gene,
	text=tumour suppressor,
	description={Genes that prevent or ``decelerate'' the process of tumorigenesis and/or tumour progression}
}

\newglossaryentry{oncogene}{
	name=Oncogene,
	text=oncogene,
	plural=oncogenes,
	description={Genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Acronym terms
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newacronym[description={$\beta$-catenin}]{bcat}{BCAT}{$\beta$-catenin}
\newacronym[description={$\gamma$-aminobutyric acid}]{gaba}{GABA}{$\gamma$-aminobutyric acid}
\newacronym[description={3-36}]{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
\newacronym[description={4-(methylnitrosamino)-1-(3-pyridyl)-1-butanone}]{nnk}{NNK}{4-(methylnitrosamino)-1-(3-pyridyl)-1-butanone}
\newacronym[description={Activation-induced cytidine deaminase}]{aid}{AID}{activation-induced cytidine deaminase}
\newacronym[description={Adenosine triphosphate}]{atp}{ATP}{adenosine triphosphate}
\newacronym[description={Adrenocorticotropic hormone}]{acth}{ACTH}{adrenocorticotropic hormone}
\newacronym[description={Agouti-related protein}]{agrp}{AGRP}{agouti-related protein}
\newacronym[description={Analysis of variance}]{anova}{ANOVA}{analysis of variance}
\newacronym[description={Apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}]{apobec}{APOBEC}{apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}
\newacronym[description={B-cell lymphoma 2}]{bcl2}{Bcl-2}{B-cell lymphoma 2}
\newacronym[description={Base-excision repair}]{ber}{BER}{base-excision repair}
\newacronym[description={Bayesian Information Criterion}]{bic}{BIC}{Bayesian Information Criterion}
\newacronym[description={Bcl-2 homology 3}]{bh3}{BH3}{Bcl-2 homology 3}
\newacronym[description={Bladder urothelial carcinoma}]{blca}{BLCA}{bladder urothelial carcinoma}
\newacronym[description={Body mass index}]{bmi}{BMI}{body mass index}
\newacronym[description={Cervical squamous cell carcinoma and endocervical adenocarcinoma}]{cesc}{CESC}{cervical squamous cell carcinoma and endocervical adenocarcinoma}
\newacronym[description={Cocaine and amphetamine related transcript}]{cart}{CART}{cocaine and amphetamine related transcript}
\newacronym[description={Colon adenocarcinoma}]{coad}{COAD}{colon adenocarcinoma}
\newacronym[description={Complimentary DNA}]{cdna}{cDNA}{complimentary DNA}
\newacronym[description={Correlation adjusted mean rank gene set test}]{camera}{CAMERA}{correlation adjusted mean rank gene set test}
\newacronym[description={Counts per million}]{cpm}{cpm}{counts per million}
\newacronym[description={Cytochrome P450}]{cyp}{CYP}{cytochrome P450}
\newacronym[description={Deoxyribonucleic acid}]{dna}{DNA}{deoxyribonucleic acid}
\newacronym[description={Differentially expressed gene}]{deg}{DEG}{differentially expressed gene}
\newacronym[description={Disability-Adjusted Life Year},longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}
\newacronym[description={Electron transport chain}]{etc}{ETC}{electron transport chain}
\newacronym[description={Endothelial-mesenchymal transition}]{emt}{EMT}{endothelial-mesenchymal transition}
\newacronym[description={Epidermal growth factor receptor}]{egfr}{EGFR}{epidermal growth factor receptor}
\newacronym[description={Estrogen receptor}]{er}{ER}{estrogen receptor}
\newacronym[description={Extracellular signal regulated kinase}]{erk}{ERK}{extracellular signal regulated kinase}
\newacronym[description={False discovery rate}]{fdr}{FDR}{false discovery rate}
\newacronym[description={Family-wise error rate}]{fwer}{FWER}{family-wise error rate}
\newacronym[description={Fat mass and obesity associated}]{fto}{FTO}{fat mass and obesity associated}
\newacronym[description={Fold change}]{fc}{FC}{fold change}
\newacronym[description={Functional class scoring}]{fcs}{FCS}{functional class scoring}
\newacronym[description={G-protein coupled receptor}]{gpcr}{GPCR}{G-protein coupled receptor}
\newacronym[description={Gap 1}]{g1}{G1}{Gap 1}
\newacronym[description={Gene Expression Omnibus}]{geo}{GEO}{Gene Expression Omnibus}
\newacronym[description={Gene Ontology}]{go}{GO}{Gene Ontology}
\newacronym[description={Gene Set Enrichment Analysis}]{gsea}{GSEA}{Gene Set Enrichment Analysis}
\newacronym[description={Genome wide association study},longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
\newacronym[description={Growth hormone secretagogue receptor}]{ghsr}{GHSR}{growth hormone secretagogue receptor}
\newacronym[description={Human epidermal growth factor receptor 2}]{her2}{HER2}{human epidermal growth factor receptor 2}
\newacronym[description={Insulin receptor substrate}]{irs}{IRS}{insulin receptor substrate}
\newacronym[description={Insulin-like growth factor binding protein}]{igfbp}{IGFBP}{insulin-like growth factor binding protein}
\newacronym[description={Insulin-like growth factor}]{igf}{IGF}{insulin-like growth factor}
\newacronym[description={Interferon-$\alpha$}]{ifna}{IFN$\alpha$}{interferon-$\alpha$}
\newacronym[description={Interferon-$\gamma$}]{ifny}{IFN$\gamma$}{interferon-$\gamma$}
\newacronym[description={Interleukin-10}]{il10}{IL-10}{interleukin-10}
\newacronym[description={Interleukin-6}]{il6}{IL-6}{interleukin-6}
\newacronym[description={International Cancer Genome Consortium}]{icgc}{ICGC}{International Cancer Genome Consortium}
\newacronym[description={Janus kinase}]{jak}{JAK}{janus kinase}
\newacronym[description={Kidney renal clear cell carcinoma}]{kirp}{KIRP}{kidney renal clear cell carcinoma}
\newacronym[description={Kyoto Encyclopedia of Genes and Genomes}]{kegg}{KEGG}{Kyoto Encyclopedia of Genes and Genomes}
\newacronym[description={Liver hepatocellular carcinoma}]{lihc}{LIHC}{liver hepatocellular carcinoma}
\newacronym[description={Lymph node}]{ln}{LN}{lymph node}
\newacronym[description={Mammalian target of rapamycin}]{mtor}{mTOR}{mammalian target of rapamycin}
\newacronym[description={Mean}]{m}{$\mu$}{mean}
\newacronym[description={Melanocortin 1 receptor}]{mc1r}{MC1R}{melanocortin 1 receptor}
\newacronym[description={Melanocortin 4 receptor}]{mc4r}{MC4R}{melanocortin 4 receptor}
\newacronym[description={Melanocyte-stimulating hormone}]{msh}{MSH}{melanocyte-stimulating hormone}
\newacronym[description={Mesenchymal-endothelial transition}]{met}{MET}{mesenchymal-endothelial transition}
\newacronym[description={Messenger RNA}]{mrna}{mRNA}{messenger RNA}
\newacronym[description={Microarray analysis suite 5.0}]{mas}{MAS5}{microarray analysis suite 5.0}
\newacronym[description={Mismatch repair}]{mmr}{MMR}{mismatch repair}
\newacronym[description={Mismatch}]{mm}{MM}{mismatch}
\newacronym[description={Neuropeptide Y1 receptor}]{nyr1}{Y1R}{neuropeptide Y1 receptor}
\newacronym[description={Neuropeptide Y}]{npy}{NPY}{neuropeptide Y}
\newacronym[description={New Zealand Breast Cancer}]{nzbc}{NZBC}{New Zealand Breast Cancer}
\newacronym[description={New Zealand}]{nz}{NZ}{New Zealand}
\newacronym[description={Next generation sequencing}]{ngs}{NGS}{next generation sequencing}
\newacronym[description={Nuclear factor $\kappa$B}]{nfkb}{NF-$\kappa$B}{nuclear factor $\kappa$B}
\newacronym[description={Nucleotide-excision repair}]{ner}{NER}{nucleotide-excision repair}
\newacronym[description={Obesity policy action}]{opa}{OPA}{obesity policy action}
\newacronym[description={Over representation analysis}]{ora}{ORA}{over representation analysis}
\newacronym[description={Pattern recognition receptor}]{prr}{PRR}{pattern recognition receptor}
\newacronym[description={Perfect match}]{pm}{PM}{perfect match}
\newacronym[description={Phosphatidylinositol 3-kinase}]{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
\newacronym[description={Phosphoenol pyruvate}]{pep}{PEP}{phosphoenol pyruvate}
\newacronym[description={Polycyclic aromatic hydrocarbon}]{pah}{PAH}{polycyclic aromatic hydrocarbon}
\newacronym[description={Pro-opiomelanocortin}]{pomc}{POMC}{pro-opiomelanocortin}
\newacronym[description={Progesterone receptor}]{pr}{PR}{progesterone receptor}
\newacronym[description={Prohormone convertase 1}]{pc1}{PC1}{prohormone convertase 1}
\newacronym[description={Reactive oxygen species}]{ros}{ROS}{reactive oxygen species}
\newacronym[description={Rectum adenocarcinoma}]{read}{READ}{rectum adenocarcinoma}
\newacronym[description={Retinoblastoma-associated}]{rb}{RB}{retinoblastoma-associated}
\newacronym[description={Ribonucleic acid}]{rna}{RNA}{ribonucleic acid}
\newacronym[description={Robust multi array}]{rma}{RMA}{robust multi array}
\newacronym[description={Signal transducer and activator of transcription 3}]{stat3}{STAT3}{signal transducer and activator of transcription 3}
\newacronym[description={Signal transducer and activator of transcription}]{stat}{STAT}{signal transducer and activator of transcription}
\newacronym[description={Single nucleotide polymorphism}]{snp}{SNP}{single nucleotide polymorphism}
\newacronym[description={Single-stranded DNA}]{ssdna}{ssDNA}{single-stranded DNA}
\newacronym[description={Singular value decomposition}]{svd}{SVD}{singular value decomposition}
\newacronym[description={Skin cutaneous melanoma}]{skcm}{SKCM}{skin cutaneous melanoma}
\newacronym[description={Standard deviation}]{sd}{$\sigma$}{standard deviation}
\newacronym[description={The Cancer Genome Atlas}]{tcga}{TCGA}{The Cancer Genome Atlas}
\newacronym[description={Thrombospondin-1}]{tsp1}{TSP-1}{thrombospondin-1}
\newacronym[description={Toll-like receptor}]{tlr}{TLR}{Toll-like receptor}
\newacronym[description={Transforming growth factor-$\beta$}]{tgfb}{TGF$\beta$}{transforming growth factor-$\beta$}
\newacronym[description={Tumour necrosis factor-$\alpha$}]{tnfa}{TNF$\alpha$}{tumour necrosis factor-$\alpha$}
\newacronym[description={Ultraviolet}]{uv}{UV}{ultraviolet}
\newacronym[description={United States of America}]{usa}{USA}{United States of America}
\newacronym[description={Uterine corpus endometrial carcinoma}]{ucec}{UCEC}{uterine corpus endometrial carcinoma}
\newacronym[description={Variance inflation factor}]{vif}{VIF}{variance inflation factor}
\newacronym[description={Vascular endothelial growth factor-A}]{vegfa}{VEGF-A}{vascular endothelial growth factor-A}
\newacronym[description={Vascular endothelial growth factor}]{vegf}{VEGF}{vascular endothelial growth factor}
\newacronym[description={Wilcoxon-Mann-Whitney}]{wmw}{WMW}{Wilcoxon-Mann-Whitney}
\newacronym[description={World Health Organization}]{who}{WHO}{World Health Organization}